---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(questionr)
library(tidyverse)
```

# Practice 1 : Vectors and sequences

## Vector creation and operation
Create the following vector 120 134 256  12
```{r}
(first_vector <- c(120, 134, 256, 12))
```

Use this vector to generate the two following vectors  
220 234 356 112
```{r}
first_vector + 100
```

240 268 512  24
```{r}
first_vector * 2
```
## Incomes
with the following vectors
```{r}
husband_income <- c(1200, 1180, 1750, 2100)
wife_income <- c(1450, 1870, 1690, 0)
household_people <- c(4, 2, 3, 2)
```

Absolute difference of income between the husband and the wife in each household (function abs)
```{r}
abs(husband_income - wife_income)
```
calculate the total income of the households
```{r}
(total_income <- husband_income + wife_income)
```
calculate the income per person in each household
```{r}
total_income / household_people
```
## Sequences
create a sequence from 1 to 10
```{r}
1:10
```

create a sequence from 0 to 100 by step of 5
```{r}
(first_seq <- seq(0, 100, 5))
```

repeat the sequence of 4. twice
```{r}
rep(first_seq, 2)
```

# Pratice 2 : Categorical variables

```{r}
education_levels <- c("Bachelor", "Master", "Ph.D")
(education <- sample(education_levels, 10, replace = TRUE))
```
what is the type of the vector
```{r}
typeof(education)
```

change the vector as a categorical variable
```{r}
(education <- factor(education))
```
display the levels and the associated integers
```{r}
levels(education)
```

```{r}
str(education)
```
change the level “Ph.D” to “PhD”
```{r}
levels(education)[3] <- "PhD"
levels(education)
```

# Data Frame
```{r}
set.seed(123) # for reproducibility
gender <- sample(c("male", "female"), 100, replace = TRUE)
age <- sample(18:65, 100, replace = TRUE)
education <- sample(c("college", "bachelor", "master", "PhD"), 100, replace = TRUE)
income <- round(rnorm(100, 50000, 10000))
children <- sample(0:4, 100, replace = TRUE)

(df <- data.frame(gender, age, education, income, children))
```

Select the age and income variables for the first 10 observations using the brackets and the $ operator
```{r}
df$age[1:10]
```

```{r}
df[1:10, c("age", "income")]
```
Select observations where the income is greater than 60,000 using the brackets and the $ operator
```{r}
df[df$income>60000,]
```

Select observations where the education level is either “college” or “master” using subset() and %in%
```{r}
df[df$education %in% c("college", "master"),]
```

```{r}
subset(df, education %in% c("college", "master"))
```

Select observations where the income is greater than 60 000 and the age is greater than 40 and display variables gender, age, income, using subset.
```{r}
subset(df, income > 60000 & age > 40)[c("gender", "age", "income")]
```

# Manipulation of functions
```{r}
temperature <- c(3.4, 4.8, 8.4, 11.4, 15.8, 19.4, 22.2, 21.6, 17.6, 13.4, 7.6, 4.4)
precipitations <- c(47.2, 44.1, 50.4, 74.9, 90.8, 75.6, 63.7, 62, 87.5, 98.6, 81.9, 55.2)
```

compute the average temperature
```{r}
mean(temperature)
```
compute the total rainfall
```{r}
sum(precipitations)
```
```{r}
python <- c(12, 16, 8, 18, 6, 10)
html <- c(14, 9, 13, 15, 17, 11)
bdd <- c(18, 11, 14, 10, 8, 12)
(notes <- cbind(python, html, bdd))
```
compute the average of each student
```{r}
rowMeans(notes)
```
compute the average in each discipline
```{r}
colMeans(notes)
```
display the quartiles of the python discipline
```{r}
quantile(notes[,1])
```
```{r}
class(notes)
```
```{r}
notes <- as.data.frame(notes)
class(notes)
```


```{r}
notes$python
```
# Data Frame function with HDV2003
```{r}
data("hdv2003")
```

```{r}
unique(hdv2003$sexe)
```
what is the mean of the variable age ?
```{r}
mean(df$age)
```
how many retired individuals are there ? (variable occup)
```{r}
nrow(hdv2003[hdv2003$occup == "Retraite", ])
```
how many retired women are there ? (variable occup and sexe)
```{r}
nrow(subset(hdv2003, occup == "Retraite" & sexe == "Femme"))
```
```{r}
nrow(hdv2003[hdv2003$occup == "Retraite" & hdv2003$sexe == "Femme", ])
```
what is the average age of retired women ?
```{r}
mean(subset(hdv2003, occup == "Retraite" & sexe == "Femme")$age)
```
```{r}
mean(hdv2003[hdv2003$occup == "Retraite" & hdv2003$sexe == "Femme", "age"])
```

# Categorical variables with forcats
with the hdv2003 dataset, recode the variable relig in order to get
```{r}
freq(hdv2003$relig)
```

```{r}
hdv2003$relig <- fct_collapse(hdv2003$relig, "Pratiquant"=c("Pratiquant regulier", "Pratiquant occasionnel"))
hdv2003$relig <- fct_recode(hdv2003$relig, "Appartenance"="Appartenance sans pratique", "NSP"="NSP ou NVPR")
freq(hdv2003$relig)
```
recode the variable nivetud to get
```{r}
freq(hdv2003$nivetud)
```
```{r}
hdv2003$nivetud <- fct_collapse(hdv2003$nivetud, 
                                "Etudes primaires"=c("A arrete ses etudes, avant la derniere annee d'etudes primaires", "Derniere annee d'etudes primaires"), "Enseignement technique ou professionel"=c("Enseignement technique ou professionnel court", "Enseignement technique ou professionnel long"))
hdv2003$nivetud <- fct_recode(hdv2003$nivetud, "Enseignement supérieur"="Enseignement superieur y compris technique superieur")
freq(hdv2003$nivetud)

```

reorder the factors of this variable to get
```{r}
hdv2003$nivetud <- fct_relevel(hdv2003$nivetud, 
                               "Enseignement supérieur", "Enseignement technique ou professionel", "2eme cycle", "1er cycle", "Etudes primaires", "N'a jamais fait d'etudes", NA)
freq(hdv2003$nivetud)
```

sort the factors of the variable relig according to the median of the variable age
```{r}
hdv2003$relig <- fct_reorder(hdv2003$relig, hdv2003$age, median)
freq(hdv2003$relig)
```

# if_else, case_when and cut
with if_else, create the variable cinema_bd which groups individuals who go to cinema and read bd. Others are put in the category “Other”. The result should be similar to
```{r}
unique(hdv2003$cinema)
```
```{r}
unique(hdv2003$lecture.bd)
```
```{r}
hdv2003$cinema_bd <- if_else(hdv2003$cinema == "Oui" & hdv2003$lecture.bd == "Oui", "Cinema and BD", "Other")
freq(hdv2003$cinema_bd)
```
```{r}
colnames(hdv2003)
```
```{r}
unique(hdv2003$trav.satisf)
```

with case_when create the variable gender_bs with the following categories : Man with more than 2 brothers and sisters, Woman with more than 2 brothers and sisters, Other. The expected result is
```{r}
hdv2003$gender_bs <- case_when(
  hdv2003$sexe == "Homme" & hdv2003$freres.soeurs > 2 ~ "Man with more than 2 brothers and sisters",
  hdv2003$sexe == "Femme" & hdv2003$freres.soeurs > 2 ~ "Woman with more than 2 brothers and sisters",
  TRUE ~ "Other"
)
freq(hdv2003$gender_bs)
```

with case_when, create a variable with the categories : Man over 30 yo, Man over 40 yo satisfied in his work, Woman practicing sport or DIY (bricolage), Other. The expected result is
```{r}
hdv2003$new_cat <- case_when(
  hdv2003$sexe == "Homme"  & hdv2003$age > 40 & hdv2003$trav.satisf == "Satisfaction" ~ "Homme de plus de 40 ans satisfait par son travail",
  hdv2003$sexe == "Homme"  & hdv2003$age > 30 ~ "Homme de plus de 30 ans",
  hdv2003$sexe == "Femme" & (hdv2003$sport == "Oui" | hdv2003$bricol == "Oui") ~ "Femme pratiquant le sport ou le bricolage",
  TRUE ~ "Other"
)
freq(hdv2003$new_cat)
```

cut the variable heures.tv to get the following frequency table
```{r}
freq(cut(hdv2003$heures.tv, breaks=c(0, 1, 2, 4, 6, 12), include.lowest = T))
```

